Multi-Agent Path Finding with Matching using Increasing Cost Tree Search

### 2021-06-16 docs/paper.tex
Finished (crude) version of complete paper

	\begin{abstract}
		Both the assignment problem and the multi-agent pathfinding problem are common problems in the fields of robotics and transportation. The joint problem of multi-agent pathfinding extended with matching between agents and goals is something that has not been studied much; few methods exist today that solve it. In this work, two types of algorithms based on the Increasing Cost Tree Search algorithm for multi-agent pathfinding are presented that can optimally solve this joint problem: exhaustive algorithms that reduce the problem to solving many MAPF problems using regular ICTS, and algorithms that search a generalized increasing cost tree. These are compared to each other experimentally on a set of grid maps, and it is shown that exhaustive methods typically outperform the generalized ICT search. An algorithm of the first type is compared to alternative algorithms based on other MAPF algorithms to put its performance into the context.
	\end{abstract}

@@

	\begin{wrapfigure}[19]{R}{.25\textwidth}
		\begin{minipage}{\linewidth}
			\centering\captionsetup[subfigure]{justification=centering}
			\includegraphics[width=\linewidth]{img/vertex-conflict}
			\subcaption{Vertex conflict}
			\label{fig:conflictsa}\par\vfill
			\includegraphics[width=0.75\linewidth]{img/edge-conflict}
			\subcaption{Edge conflict}
			\label{fig:conflictsb}
		\end{minipage}
		\vspace{-10pt}
		\caption{The two types of conflict (from Fig. 1 in \cite{stern2019})}\label{fig:conflicts}
	\end{wrapfigure}
	This work is the outcome of a search for algorithms for optimally solving MAPFM based on ICTS as an alternative to the approach outlined above\footnote{To the knowledge of the author, to date, no such algorithm is described in the literature}, which took place as part of a project in which  other algorithms addition to ICTS were taken as starting points for MAPFM algorithms. Two classes of ICTS-based algorithms were found: one based on exhaustive enumeration of matchings using a variant of ICTS as subroutine, the other based on modifications to the ICT structure and its search. One exhaustive ICTS algorithm is compared to algorithms based on M*\cite{wagner2011}, A*+OD+ID\cite{standley2010}, CBM\cite{ma2016} and EPEA*\cite{goldenberg2014} respectively to put ICTS methods for MAPFM into a broader context.

@@

	In practice, joint MDDs are not explicitly constructed but instead, a depth-first search is used to search joint MDDs. This has the benefit of finding a path to the bottom of the joint MDD fast if one exists, while not constructing more of the joint MDD (which can get rather large) than is necessary for finding a solution.
	\subsubsection{ICTS with Independence Detection}
	Like other MAPF algorithms, ICTS can be embedded in the Independence Detection (ID) framework introduced by Standley in \cite{standley2010}. ID starts by assuming each agent is a group that can be assigned an optimal path independent of other groups. If group paths do conflict, they become one group that is planned together using a MAPF algorithm, in this case, ICTS, until no conflicts between groups occur. The goal of ID is to minimize $k'$, the size of the largest group that has to be planned. Optionally, a conflict avoidance table (CAT) can be maintained for use during the planning of groups. This records the paths planned by other groups and can facilitate tie-breaking in search that minimizes the number of conflicts; the CAT is omitted from Algorithm \ref{id-algo} for brevity but was implemented and used in experiments.
	\begin{algorithm}
		\begin{algorithmic}[1]
			\State \text{Assign each agent to a singleton group}
			\State \text{plan a path for each group}
			\Repeat
			\State \text{simulate execution of all paths until a conflict occurs}
			\State \text{merge two conflicting groups into a single group}
			\State \text{cooperatively plan new group}
			\Until{\text{no conflicts occur}}
			\State $solution\gets\text{paths of all groups combined}$
			\State \textbf{return solution}
		\end{algorithmic}
		\caption{Simple Independence Detection by Standley} 
		\label{id-algo}
	\end{algorithm}
	\subsubsection{Pruning in ICTS}
	\label{pruning}
	In \cite{sharon2011}, in addition to the basic two-level solver, three pruning methods to speed up search are discussed, two of which were used in this work. All work by taking $j < k$ and applying the described method to each possible combination of $j$ out of $k$ agents. If no solution exists for any such combination, the full $k$-agent low-level search is not triggered. The two basic pruning methods are as follows.
	\begin{enumerate}
		\item Simple pruning: the $j$-agent MDD is searched using DFS for a solution.
		\item Enhanced pruning: the $j$-agent MDD is searched using BFS and when children are generated, for each agent, all children that do not appear in any unified node are removed from the MDD. In \ref{fig:mdds}, the $C$ node of $MDD_1$ would be removed as it cannot be unified with any node of $MDD_2$ at that level. 
	\end{enumerate}
	In addition, repeated enhanced pairwise pruning is described in which the enhanced pruning procedure is repeated for all combinations until a fixpoint is reached, which is when no child node is removed from any of the agent MDDs in an iteration. This was not used in this work.

	The full ICT-search algorithm including a pruning step is described by Algorithm \ref{icts-algo}. It is reproduced from \cite{sharon2011} for completeness. As noted before, this procedure could be used for planning groups in Algorithm \ref{id-algo}. In this work, triple-pruning was used and enhanced triple-pruning was taken as a base-line, as the conclusion in \cite{sharon2011} is that this generally outperforms other variants.
	\begin{algorithm}
		\begin{algorithmic}[1]
			\Procedure{ICT-search}{MAPF instance}
			\State \text{Construct ICT root}
			\ForAll{\text{ICT nodes }$(c_1,\ldots,c_k)$\text{ in breadth-first order }} 
			\ForAll{\text{agents} $a_i$} 
			\text{Construct} $MDD_i^{c_i}$
			\EndFor
			\ForAll{$j$-\text{agent combinations} $(i_1,\ldots,i_j)$} 
			\State \text{Run node-pruning procedure with MDDs }$MDD_{i_1}^{c_{i_1}},\ldots,MDD_{i_j}^{c_{i_j}}$
			\If{\text{no solution was found for a combination in pruning}}
			\State \textbf{Break}\text{ and continue with next ICT node}
			\EndIf
			\EndFor
			\State \text{Search }$k$\text{-agent MDD space}
			\If{\text{solution was found}}
			\State \textbf{return solution}
			\EndIf
			\EndFor

			\EndProcedure
		\end{algorithmic}
		\caption{Increasing Cost Tree Search}
		\label{icts-algo}
	\end{algorithm}
	\subsection{Exhaustive ICTS}
	\label{exhaustive}
	For $K$ teams $t_1,\ldots,t_K$ of size $k_i$, there are in total $M = \prod_{i} k_i!$ matchings from team agents to team goals. Let $C_1$ through $C_M$ denote the optimal SIC costs of each MAPF instance corresponding to the instance, with $C_m = \infty$ for unfeasible $m$. In order to find the optimal solution, $m$ with lowest $C_m$ has to be found, which requires all $M$ to be solved or found to be unfeasible, meaning that $\mathcal{O}(M)$ MAPF instances have to be solved. 
	\subsubsection{Bounded search}
	Once the first solution is found with cost $C$, this can be used in the next search as an (exclusive) upper-bound $B$ for the SIC of nodes that have to be searched. Formally, let $B_0 = \infty$ represent the initial SIC bound of the ICT search. After solving each $m$ with solution cost $C_m$, let $B_{m+1} = \min(B_m,C_m)$. By successively updating this $B$ in this manner, as more matchings are solved, the effort needed to solve the remaining matchings can be expected to be reduced significantly. With $C_m$ denoting the optimal SIC for matching $m$, this effort can be expressed as	
	\[\Delta_{m,bounded} = \min(B_m,C_m) - C^*_m = B_{m+1} - C^*_m\]
	This is proportional to the effort due to the complexity of the ICTS search for matching $m$, $\mathcal{O}(k^{\Delta_{m,bounded}})$.
	If $B_m < C^*_m$, there is no need to consider $m$ and $\Delta_{m,bounded} = 0$.
	Using this, the number of ICT nodes searched in total becomes $\mathcal{O}(\sum_m k^{\Delta_{m,bounded}})$.
	\subsubsection{Ordered enumeration}
	\label{ordered-enum}
	In an effort to reduce the total effort as much as possible, the matchings $m$ could be enumerated and solved in increasing order of $C^*_m$. In this case, the effort per matching is monotonically non-increasing by Property \ref{monotonicity}. This induces a case in which this enumeration minimizes the number of nodes searched in total, $\Delta_m$ (Property \ref{cond-monoton-optim}). However, cases in which $\Delta_m$ of the first $m$'s searched is much higher could also be constructed, in which case this approach might cause more nodes to be processed in total than alternative methods.

@@

	\subsubsection{MDD generation for matching}
	To generate MDDs for MAPFM, each agent also has to consider not one goal but all matching goals. By following a BFS-based approach, with all team goals being goals in the search and in turn endpoints from which, using the time-expanded graph, the MDD was generated that is used in the low-level search, agent MDDs can be made to represent all possible paths of a set length to any matching goal.
	\begin{wrapfigure}{R}{.35\textwidth}
		\centering
		\includegraphics[width=\linewidth]{img/path}
		\caption{Pathological case for ICTS-m}
		\label{fig:path}
	\end{wrapfigure}
	\subsubsection{Pathological case}
	\label{pathology}
	As a consequence of the $\min$ operation in the root definition, a pathological case can be derived. Consider a team of 4 agents $a_1^1$ through $a_4^1$. Let them be arranged in a configuration as depicted in Figure \ref{fig:path}. All four agents match the colour of the centre goal and hence for each $i$, $c^*_i = 1$, so that the root is $(1,1,1,1)$ with SIC $4$. The optimal solution however is, numbering the agents clockwise, $(1,3,3,3)$ with SIC $10$, meaning that $\Delta = 6$. It can be seen that instances of this form with arbitrarily large $\Delta$ can be constructed. As ICTS-m being a generalization of ICTS is $\mathcal{O}(k^\Delta)$, even just extending the 'limbs' of the instance to yield an ICT root of $(1,9,9,9)$ with $\Delta = 24$ will result in an unfathomably large number of nodes being evaluated assuming no pruning or other modifications to the base algorithm. [TODO: reduce case to just 2 agents?]
	%\subsubsection{Inverted MDD search}
	%The uniqueness of goals guarantees the existence of a goal node without vertex conflicts in searching MDDs in MAPF. In MAPFM, this is not the case and causes joint MDDs to be searched fully without finding a result. In addition, even if there exists a goal node, many of the paths represented by the joint MDD lead to conflicting maximal-depth nodes.

	%To avoid this, the MDD search could be changed to start at the maximal depth and search for a way towards the start node, which by start uniqueness remains without vertex conflicts.
	%Two strategies could be used to avoid this: firstly, the MDD search could be changed to start at the maximal depth and search for a way towards the start (which by start uniqueness is always non-conflicting); secondly, a bipartite matching algorithm could be employed to see if any matching exists between agents and goals they can reach. An advantage of the former strategy is that it will, in particular with larger team sizes, reduce the number of maximal depth nodes drastically, so that the subsequent search is from few nodes to one goal. By only running a matching algorithm to skip regular MDD search, there is no such reduction so from one starting position paths will be searched to potentially an exponential number of goal nodes, of which many will end up being illegal. 
	%[TODO: just focus on explaining why inverted MDD search was/could be used to solve this? Or just toss this out for lack of time/space. Why didn't I come up with this idea earlier...]
	\subsection{Optimizations}
	To improve both types of algorithms, some optimizations were found that can be used to speed up the search. These are here motivated. 
	\subsubsection{Bounds in ID}
	Information gained about solution costs of independent groups can be used to establish a lower bound on solution SIC as per Property \ref{property:id-lower}, and in bounded search the upper bound can be further constrained (Property \ref{property:id-upper}).
	\begin{property}[Lower bound]
		When merging groups $g_i,g_j$ with solution costs $C_i,C_j$ to yield group $g_k$, for any solution for this group with cost $C_k$
		\[C_k\geq C_i + C_j\]
		\label{property:id-lower}
	\end{property}
	\begin{proof}
		Assume there would be a solution with $C_k < C_i + C_j$. Taking $C_k^i$ to denote the SIC of agents that used to make up $g_i$ in the solution to $g_k$, without loss of generality $C_k^i < C_i$, which contradicts the optimality of the solution with cost $C_i$ for $g_i$ independent of $g_j$.
	\end{proof}
	\begin{property}[Reduced upper bound]
		In bounded exhaustive search, the sum of costs of groups not being merged can be subtracted from bound $B$ without losing optimality.
		\label{property:id-upper}
	\end{property}
	\subsubsection{Child pruning}
	In the original ICTS formulation all $k$-children are generated in every case because pruning only controls skipping the low-level search, assumed to be the most expensive operation. However, in simple pruning in any case, when there is a conflict between two or more agents at a given ICT node $(c_1,\ldots,c_k)$, in any solution, one or more of the path costs corresponding to these conflicting agents will have to be incremented. This can be explioted by pruning the node's children that do not correspond to the conflicting agents, of which there are often just 2 or 3, drastically improving the (average) branching factor as $k$ increases. By Property \ref{property:ec-incompat}, child pruning is not compatible with enhanced pruning.
	\begin{property}[Incompatibility with enhanced pruning]
		Enhanced pruning combined with child pruning (hereafter E+C pruning) is suboptimal due to the \textit{cascading effect} described in \cite{sharon2011}.
		\label{property:ec-incompat}
	\end{property}

	\begin{figure}[H]
	\centering
	\begin{subfigure}{0.40\textwidth}
		\centering
		\includegraphics[width=\linewidth]{img/counter-example-ep-2}
		\caption{Instance}
		\label{fig:counter-example:a}
	\end{subfigure}
	\begin{subfigure}{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{img/counter-example-tree}
		\caption{Pruned ICT that is traversed}
		\label{fig:counter-example:b}
	\end{subfigure}
	\caption{Counter-example for E+C pruning}
	\label{fig:counter-example}
\end{figure}
	\begin{proof}
		Consider Figure \ref{fig:counter-example:a} with a red agent $r$, green agent $g$ and black agent $b$. It can be seen that $(3,8,5)$ with SIC $16$ is an optimal solution for this instance, which corresponds to $r$ moving aside, allowing $g$ to pass.  Say enhanced pruning with pairs is applied to these three agents at $(c^*_r,c^*_g,c^*_b) = (0,8,5)$. Let the enhanced pruning order be $(r,g)$, $(r,b)$, $(g,b)$ without loss of generality. $(r,g)$ is solvable but will leave for $g$ out of the two paths of length $8$ only the clock-wise path. $(r,b)$ will not change either MDD in this node. When $(g,b)$ is evaluated, it is found that the remaining clockwise path for $g$ and $b$'s path are incompatible (edge conflict) so $(0,9,5)$ and $(0,8,6)$ are generated. By repeating this order of enhanced pruning with child pruning, the ICT shown in \ref{fig:counter-example:b} is traversed. Notice that none of the nodes with SIC 16 in this ICT are a solution, meaning that E+C pruning will return a suboptimal solution to the instance in \ref{fig:counter-example:a}, making it suboptimal.
	\end{proof}

	\section{Results}
	\label{experiments}
	To compare the algorithms and variants described above and put them into the context of alternative MAPFM algorithms, two types of experiments were performed. In the first, ICTS-based algorithms discussed in this work are compared; in the second, the approximate ordered enumeration variant is compared to other algorithms for MAPFM developed within the peer group\footnote{I want to thank Jonathan Dönszelmann, who worked on M*-based algorithms, for setting up this comparison[TODO: move to thanks note at end]}. The values compared are the fraction of instances solved, and the average runtime per instance. It should be noticed that in the latter, runtimes converge to the 120s timeout used as fewer instances are solved; therefore, while the times are included for completeness at the end of this work, the fraction of instances solved is guiding in the analysis of results.
		\newgeometry{left=1cm,right=1cm,top=0.1cm,bottom=0.1cm}
	\begin{figure*}[t]
		\centering
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/25-1-p}
			\caption{25, 1 team}
			\label{fig:i-25-1-p}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/25-3-p}
			\caption{25, 3 teams}
			\label{fig:i-25-3-p}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/75-1-p}
			\caption{75, 1 team}
			\label{fig:i-75-1-p}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/75-3-p}
			\caption{75, 3 teams}
			\label{fig:i-75-3-p}
		\end{subfigure}
		\caption{Fraction solved out of 50 maps on two types of 20x20 grids with a 120s timeout}
		\label{fig:i-probs}
	\end{figure*}
	\begin{figure*}[b]
		\centering
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/25-1-p}
			\caption{25, 1 team}
			\label{fig:r-25-1-p}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/25-3-p}
			\caption{25, 3 teams}
			\label{fig:r-25-3-p}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/75-1-p}
			\caption{75, 1 team}
			\label{fig:r-75-1-p}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/75-3-p}
			\caption{75, 3 teams}
			\label{fig:75-3-p}
		\end{subfigure}
		\caption{Fraction solved out of 50 maps on two types of 20x20 grids with a 120s timeout}
		\label{fig:r-probs}
	\end{figure*}

>>>>

	\begin{abstract}
		Both the assignment problem and the multi-agent pathfinding problem are common problems in the fields of robotics and transportation. The joint problem of multi-agent pathfinding extended with matching between agents and goals is something that has not been studied much; few methods exist today that solve it. In this work, two types of algorithms based on the Increasing Cost Tree Search algorithm for multi-agent pathfinding are presented that can optimally solve this joint problem: exhaustive algorithms that reduce the problem to solving many MAPF problems using regular Increasing Cost Tree Search, and algorithms that search a generalized increasing cost tree. These are compared to each other experimentally on a set of grid maps, and it is shown that exhaustive methods typically outperform the generalized ICT search. Lastly, an exhaustive algorithm is compared to alternative algorithms based on other multi-agent pathfinding approaches to put its performance into the broader context of algorithms for multi-agent pathfinding with matching.
	\end{abstract}

@@

	\begin{wrapfigure}[19]{R}{.25\textwidth}
		\begin{minipage}{\linewidth}
			\centering\captionsetup[subfigure]{justification=centering}
			\includegraphics[width=\linewidth]{img/vertex-conflict}
			\subcaption{Vertex conflict}
			\label{fig:conflictsa}\par\vfill
			\includegraphics[width=0.75\linewidth]{img/edge-conflict}
			\subcaption{Edge conflict}
			\label{fig:conflictsb}
		\end{minipage}
		\vspace{-10pt}
		\caption{The two types of conflict (from Fig. 1 in \cite{stern2019})}\label{fig:conflicts}
	\end{wrapfigure}
	This work is the outcome of a search for algorithms for optimally solving MAPFM based on ICTS as an alternative to the approach outlined above\footnote{To the knowledge of the author, to date, no such algorithm is described in the literature}, which is part of a project in which  other algorithms addition to ICTS were taken as starting points for MAPFM algorithms. Two classes of ICTS-based algorithms were found: one based on exhaustive enumeration of matchings using a variant of ICTS as subroutine, the other based on modifications to the ICT structure and its search. One exhaustive ICTS algorithm is compared to algorithms based on CBM\cite{ma2016}, M*\cite{wagner2011}, A*+OD+ID\cite{standley2010} and EPEA*\cite{goldenberg2014} respectively to put ICTS methods for MAPFM into a broader context.

@@

	In practice, joint MDDs are not explicitly constructed but instead, a depth-first search is used to search joint MDDs. This has the benefit of finding a path to the bottom of the joint MDD fast if one exists, while not constructing more of the joint MDD (which can get rather large) than is necessary for finding a solution.
	\subsubsection{ICTS with Independence Detection}
	Like other MAPF algorithms, ICTS can be embedded in the Independence Detection (ID) framework introduced by Standley in \cite{standley2010}. ID starts by assuming each agent is a group that can be assigned an optimal path independent of other groups. If group paths do conflict, they become one group that is planned together using a MAPF algorithm, in this case, ICTS, until no conflicts between groups occur. The goal of ID is to minimize $k'$, the size of the largest group that has to be planned. Optionally, a conflict avoidance table (CAT) can be maintained for use during the planning of groups. This records the paths planned by other groups and can facilitate tie-breaking in search that minimizes the number of conflicts; the CAT is omitted from Algorithm \ref{id-algo} for brevity but was implemented and used in experiments.
	\begin{algorithm}
		\begin{algorithmic}[1]
			\State \text{Assign each agent to a singleton group}
			\State \text{plan a path for each group}
			\Repeat
			\State \text{simulate execution of all paths until a conflict occurs}
			\State \text{merge two conflicting groups into a single group}
			\State \text{cooperatively plan new group}
			\Until{\text{no conflicts occur}}
			\State $solution\gets\text{paths of all groups combined}$
			\State \textbf{return solution}
		\end{algorithmic}
		\caption{Simple Independence Detection \cite{standley2010}} 
		\label{id-algo}
	\end{algorithm}
	\subsubsection{Pruning in ICTS}
	\label{pruning}
	In \cite{sharon2011}, in addition to the basic two-level solver, three pruning methods to speed up search are discussed, two of which were used in this work. All work by taking $j < k$ and applying the described method to each possible combination of $j$ out of $k$ agents. If no solution exists for any such combination, the full $k$-agent low-level search is not triggered. The two basic pruning methods are as follows.
	\begin{enumerate}
		\item Simple pruning: the $j$-agent MDD is searched using DFS for a solution.
		\item Enhanced pruning: the $j$-agent MDD is searched using BFS and when children are generated, for each agent, all children that do not appear in any unified node are removed from the MDD. In \ref{fig:mdds}, the $C$ node of $MDD_1$ would be removed as it cannot be unified with any node of $MDD_2$ at that level. 
	\end{enumerate}
	In addition, repeated enhanced pairwise pruning is described in which the enhanced pruning procedure is repeated for all combinations until a fixpoint is reached, which is when no child node is removed from any of the agent MDDs in an iteration. This was not used in this work.

	The full ICT-search algorithm including a pruning step is described by Algorithm \ref{icts-algo}. It is reproduced from \cite{sharon2011} for completeness. As noted before, this procedure can be used for planning groups in Algorithm \ref{id-algo}. In this work, triple-pruning was used and enhanced triple-pruning was taken as a base-line, as the conclusion in \cite{sharon2011} is that this often outperforms other variants.
	\begin{algorithm}
		\begin{algorithmic}[1]
			\Procedure{ICT-search}{MAPF instance}
			\State \text{Construct ICT root}
			\ForAll{\text{ICT nodes }$(c_1,\ldots,c_k)$\text{ in breadth-first order }} 
			\ForAll{\text{agents} $a_i$} 
			\text{Construct} $MDD_i^{c_i}$
			\EndFor
			\ForAll{$j$-\text{agent combinations} $(i_1,\ldots,i_j)$} 
			\State \text{Run node-pruning procedure with MDDs }$MDD_{i_1}^{c_{i_1}},\ldots,MDD_{i_j}^{c_{i_j}}$
			\If{\text{no solution was found for a combination in pruning}}
			\State \textbf{Break}\text{ and continue with next ICT node}
			\EndIf
			\EndFor
			\State \text{Search }$k$\text{-agent MDD space}
			\If{\text{solution was found}}
			\State \textbf{return solution}
			\EndIf
			\EndFor

			\EndProcedure
		\end{algorithmic}
		\caption{Increasing Cost Tree Search}
		\label{icts-algo}
	\end{algorithm}
	\subsection{Exhaustive ICTS}
	\label{exhaustive}
	For $K$ teams $t_1,\ldots,t_K$ of size $k_i$, there are in total $M = \prod_{i} k_i!$ matchings from team agents to team goals. Let $C_1$ through $C_M$ denote the optimal SIC costs of each MAPF instance corresponding to the instance, with $C_m = \infty$ for unfeasible $m$. In order to find the optimal solution, $m$ with lowest $C_m$ has to be found, which requires all $M$ to be solved or found to be unfeasible, meaning that $\mathcal{O}(M)$ MAPF instances have to be solved. 
	\subsubsection{Bounded search}
	Once the first solution is found with cost $C$, this can be used in the next search as an (exclusive) upper-bound $B$ for the SIC of nodes that have to be searched. Formally, let $B_0 = \infty$ represent the initial SIC bound of the ICT search. After solving each $m$ with solution cost $C_m$, let $B_{m+1} = \min(B_m,C_m)$. By successively updating this $B$ in this manner, as more matchings are solved, the effort needed to solve the remaining matchings can be expected to be reduced significantly. With $C_m$ denoting the optimal SIC for matching $m$, this effort can be expressed as	
	\[\Delta_{m,bounded} = \min(B_m,C_m) - C^*_m = B_{m+1} - C^*_m\]
	This is proportional to the effort due to the complexity of the ICTS search for matching $m$, $\mathcal{O}(k^{\Delta_{m,bounded}})$.
	If $B_m < C^*_m$, only the root node will be generated to determine this so $\Delta_{m,bounded} = 0$.
	Using this, the number of ICT nodes searched in total becomes $\mathcal{O}(\sum_m k^{\Delta_{m,bounded}})$.
	\subsubsection{Ordered enumeration}
	\label{ordered-enum}
	In an effort to reduce the total effort as much as possible, the matchings $m$ can be enumerated and solved in increasing order of $C^*_m$. In this case, the effort per matching is monotonically non-increasing by Property \ref{monotonicity}. This induces a case in which this enumeration minimizes the number of nodes searched in total (Property \ref{cond-monoton-optim}). However, cases in which $\Delta_m$ of the first $m$'s searched is much higher could also be constructed, in which case this approach might cause more nodes to be processed in total than alternative methods. Therefore, this ordered enumeration remains a heuristic improvement that works well in practice, as shown in Section \ref{experiments}.

@@

	\subsubsection{MDD generation for matching}
	To generate MDDs for MAPFM, each agent also has to consider not one goal but all matching goals. By following a BFS-based approach, with all team goals being goals in the search and in turn endpoints from which, using the time-expanded graph, the MDD was generated that is used in the low-level search, agent MDDs can be made to represent all possible paths of a set length to any matching goal.
	\begin{wrapfigure}{R}{.35\textwidth}
		\centering
		\includegraphics[width=\linewidth]{img/path}
		\caption{Pathological case for ICTS-m}
		\label{fig:path}
	\end{wrapfigure}
	\subsubsection{Pathological case}
	\label{pathology}
	As a consequence of the $\min$ operation in the root definition, a pathological case can be derived. Consider a team of 4 agents $a_1^1$ through $a_4^1$. Let them be arranged in a configuration as depicted in Figure \ref{fig:path}. All four agents match the colour of the centre goal and hence for each $i$, $c^*_i = 1$, so that the root is $(1,1,1,1)$ with SIC $4$. The optimal solution however is, numbering the agents clockwise, $(1,3,3,3)$ with SIC $10$, meaning that $\Delta = 6$. It can be seen that instances of this form with arbitrarily large $\Delta$ can be constructed. As ICTS-m being a generalization of ICTS is $\mathcal{O}(k^\Delta)$, even just extending the 'limbs' of the instance to yield an ICT root of $(1,9,9,9)$ with $\Delta = 24$ will result in an unfathomably large number of nodes being evaluated assuming no pruning or other modifications to the base algorithm.
	%\subsubsection{Inverted MDD search}
	%The uniqueness of goals guarantees the existence of a goal node without vertex conflicts in searching MDDs in MAPF. In MAPFM, this is not the case and causes joint MDDs to be searched fully without finding a result. In addition, even if there exists a goal node, many of the paths represented by the joint MDD lead to conflicting maximal-depth nodes.

	%To avoid this, the MDD search could be changed to start at the maximal depth and search for a way towards the start node, which by start uniqueness remains without vertex conflicts.
	%Two strategies could be used to avoid this: firstly, the MDD search could be changed to start at the maximal depth and search for a way towards the start (which by start uniqueness is always non-conflicting); secondly, a bipartite matching algorithm could be employed to see if any matching exists between agents and goals they can reach. An advantage of the former strategy is that it will, in particular with larger team sizes, reduce the number of maximal depth nodes drastically, so that the subsequent search is from few nodes to one goal. By only running a matching algorithm to skip regular MDD search, there is no such reduction so from one starting position paths will be searched to potentially an exponential number of goal nodes, of which many will end up being illegal. 
	%[TODO: just focus on explaining why inverted MDD search was/could be used to solve this? Or just toss this out for lack of time/space. Why didn't I come up with this idea earlier...]
	\subsection{Optimizations}
	To improve both types of algorithms, some optimizations were found that can be used to speed up the search. These are here motivated. 
	\subsubsection{Bounds in ID}
	Information gained about solution costs of independent groups can be used to establish a lower bound on solution SIC as per Property \ref{property:id-lower}. In bounded search the upper bound can be further constrained by Property \ref{property:id-upper}.
	\begin{property}[Lower bound]
		When merging groups $g_i,g_j$ with solution costs $C_i,C_j$ to yield group $g_k$, for any solution for this group with cost $C_k$
		\[C_k\geq C_i + C_j\]
		\label{property:id-lower}
	\end{property}
	\begin{proof}
		Assume there would be a solution with $C_k < C_i + C_j$. Taking $C_k^i$ to denote the SIC of agents that used to make up $g_i$ in the solution to $g_k$, without loss of generality $C_k^i < C_i$, which contradicts the optimality of the solution with cost $C_i$ for $g_i$ independent of $g_j$.
	\end{proof}
	\begin{property}[Reduced upper bound]
		In bounded exhaustive search, the sum of costs of groups not being merged can be subtracted from bound $B$ without losing optimality.
		\label{property:id-upper}
	\end{property}
	\subsubsection{Child pruning}
	In the original ICTS formulation all $k$-children are generated in every case because pruning only controls skipping the low-level search, assumed to be the most expensive operation. However, in simple pruning in any case, when there is a conflict between two or more agents at a given ICT node $(c_1,\ldots,c_k)$, in any solution, one or more of the path costs corresponding to these conflicting agents will have to be incremented. This can be explioted by pruning the node's children that do not correspond to the conflicting agents, of which there are often just 2 or 3, drastically improving the (average) branching factor as $k$ increases. By Property \ref{property:ec-incompat}, child pruning is not compatible with enhanced pruning.
	\begin{property}[Incompatibility with enhanced pruning]
		Enhanced pruning combined with child pruning is suboptimal due to the \textit{cascading effect} described in \cite{sharon2011}.
		\label{property:ec-incompat}
	\end{property}

	\begin{figure}[H]
	\centering
	\begin{subfigure}{0.40\textwidth}
		\centering
		\includegraphics[width=\linewidth]{img/counter-example-ep-2}
		\caption{Instance}
		\label{fig:counter-example:a}
	\end{subfigure}
	\begin{subfigure}{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{img/counter-example-tree}
		\caption{Pruned ICT that is traversed}
		\label{fig:counter-example:b}
	\end{subfigure}
	\caption{Counter-example for enhanced pruning combined with child pruning}
	\label{fig:counter-example}
\end{figure}
	\begin{proof}
		Consider Figure \ref{fig:counter-example:a} with a red agent $r$, green agent $g$ and black agent $b$. It can be seen that $(3,8,5)$ with SIC $16$ is an optimal solution for this instance, which corresponds to $r$ moving aside, allowing $g$ to pass.  Say enhanced pruning with pairs is applied to these three agents at $(c^*_r,c^*_g,c^*_b) = (0,8,5)$. Let the enhanced pruning order be $(r,g)$, $(r,b)$, $(g,b)$ without loss of generality. $(r,g)$ is solvable but will leave for $g$ out of the two paths of length $8$ only the clock-wise path. $(r,b)$ will not change either MDD in this node. When $(g,b)$ is evaluated, it is found that the remaining clockwise path for $g$ and $b$'s path are incompatible (edge conflict) so $(0,9,5)$ and $(0,8,6)$ are generated. By repeating this order of enhanced pruning with child pruning, the ICT shown in \ref{fig:counter-example:b} is traversed. Notice that none of the nodes with SIC 16 in this ICT are a solution, meaning that enhanced pruning combined with child pruning will return a suboptimal solution to the instance in \ref{fig:counter-example:a}, making this pruning strategy suboptimal.
	\end{proof}

	\section{Experimental Results}
	\label{experiments}
	To compare the algorithms and variants described above with each other and with alternative MAPFM algorithms, two types of experiments were performed. In the first, ICTS-based algorithms discussed in this work are compared; in the second, the approximate ordered enumeration variant is compared to other algorithms for MAPFM developed within the peer group. This comparison data originates in \cite{donszelmann2021}. The values compared are the fraction of instances solved, and the average runtime per instance. It should be noticed that in the latter, the recorded runtimes converge to the 120s timeout used as fewer instances are solved; therefore, while the times are included for completeness at the end of this work (Figure \ref{fig:i-times} and \ref{fig:r-times} for experiments 1 and 2), the fraction of instances solved is guiding in the analysis of results.
	\newgeometry{left=1cm,right=1cm,top=0.1cm,bottom=0.1cm}
	\begin{figure*}[t]
		\centering
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/25-1-p}
			\caption{25\% wall, 1 team}
			\label{fig:i-25-1-p}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/25-3-p}
			\caption{25\% wall, 3 teams}
			\label{fig:i-25-3-p}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/75-1-p}
			\caption{75\% wall, 1 team}
			\label{fig:i-75-1-p}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/icts-comparison/75-3-p}
			\caption{75\% wall, 3 teams}
			\label{fig:i-75-3-p}
		\end{subfigure}
		\caption{Fraction solved out of 50 maps by different ICTS-based algorithms}
		\label{fig:i-probs}
	\end{figure*}
	\begin{figure*}[b]
		\centering
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/25-1-p}
			\caption{25\% wall, 1 team}
			\label{fig:r-25-1-p}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/25-3-p}
			\caption{25\% wall, 3 teams}
			\label{fig:r-25-3-p}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/75-1-p}
			\caption{75\% wall, 1 team}
			\label{fig:r-75-1-p}
		\end{subfigure}
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/results/relative-comparison/75-3-p}
			\caption{75\% wall, 3 teams}
			\label{fig:r-75-3-p}
		\end{subfigure}
		\caption{Fraction solved out of 200 maps by MAPFM algorithms}
		\label{fig:r-probs}
	\end{figure*}
